---
title: "Zhao Ping HW2"
author: "Ping Zhao"
date: "2/10/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

###(a) Gradient Function
```{r}
#x=(x1,x2,x3,x4), and t os the parameter (theta)
grad <- function (x, t) {
  dl = x[1]/(2+t)-(x[2]+x[3])/(1-t)+x[4]/t
}


```




###(b)Stop your algorithm when either of the following criteria is satisfied:

```{r}
secant <- function(maxit, x, t0, t1, tolerr, tolgrad){
  tstar = -1657/7680+sqrt(3728689)/7680
  digits = -log10(abs(t1 - tstar))
  it=0
  stop=0
  
  theta=c()
  converage.rate=c()
  digit=c()
  iteration=c()
  modified.relerr=c()
  Gradients=c()
  while (it < maxit & stop==0){
      it = it+1
      dl.1 = grad(x,t1)
      dl.0 = grad(x,t0)
      
      tnew = t1-dl.1*(t1-t0)/(dl.1-dl.0)
  
      
      rate = abs(tnew - tstar)/abs(t1 - tstar) #check convergent rate
      digits = -log10(abs(tnew - tstar)/abs(tstar))
      mod.relerr = abs(t1 - tnew)/max(1,abs(tnew))
      dl.new <-grad(x,tnew)
      
      if (mod.relerr < tolerr & abs(dl.new) < tolgrad) stop=1 # Stop iteration condition
      #print(c(it, t, rate,digits))
      #print(sprintf('it = %2.0f   teta = %12.12f    Rate = %4.2f     digits = %2.1f',it,tnew,rate,digits))
      #print(sprintf('          relerr = %4.1e,        grad = %4.1e',relerr,dl))
      t0 = t1 # Update t0
      t1 = tnew # Update and return
      
      theta[it]<- tnew
      converage.rate[it]<- rate
      digit[it]<-digits
      iteration[it]<- it
      modified.relerr[it] <-mod.relerr
      Gradients[it]<-dl.1
    }
  conver.info<-as.data.frame(cbind(iteration, theta, converage.rate, digit, modified.relerr, Gradients))
  return(conver.info)
}

a=c()


```




### (c) &(d)

$\theta^{(0)}=.02$ and $\theta^{(1)}=.01$, and use tolerr = 1e-6 and tolgrad=1e-9

The iteration information is in matrix m. A larger matrix m.index contains information of starting points.
```{r}

d <- c(1997,907,904,32)

#secant(20,x=d,0.02,0.01,1e-6,1e-9 )


m<-secant(20,x=d,0.02,0.01,1e-6,1e-9)
m
```

The matrix above gives iteration information about the secant method.  



However, I need to add information of the starting point $\theta_0$ and $\theta_1$ to the convergence matrix. As we are applying secant method, the $\theta_2$ is from the first iteration. (I don't think this index system is good.) I added an column of $\theta$ index to the matrix.

```{r}


Index.Theta<-seq(0,nrow(m)+1)
tstar = -1657/7680+sqrt(3728689)/7680

#mod.relerr.0 = abs(0.02 - tnew)/max(1,abs(tnew))
mod.relerr.1 = abs(0.01 - 0.02)/max(1,abs(0.01))

rate.t1 = abs(0.01 - tstar)/abs(0.02 - tstar)

digits.1 = -log10(abs(0.01 - tstar)/abs(tstar))
digits.0 = -log10(abs(0.02 - tstar)/abs(tstar))

t0.info<-c(NA,0.02,NA,digits.0, NA, grad(d,0.02))
t1.info<-c(NA,0.01,rate.t1,digits.1,mod.relerr.1,grad(d,0.01))

m.starting<-rbind(t0.info,t1.info,m)
m.index<-cbind(Index.Theta,m.starting)

m.index

```



###(e)
```{r}

rate.sq<-m$converage.rate^2

plot(m$converage.rate, ylab = "Rate", main = "Converagence Rate", pch=15)

abline(h=0, col=2)


```

As we can see from the rate plot above, the rate is less than 1 and going to 0. Thus, it is supper-linear converagence.



###(f)

$\theta^{(0)}=0.5$ and $\theta^{(1)}=0.01$

```{r}
m.f<-secant(1000,x=d,0.5,0.01,1e-6,1e-6)
plot(m.f$theta)
```

From the result above, we can see that the secant method doesn't converage in this case.

## Problem 2
(a) log likelihood function

$$
l=-20*log2\pi+\Sigma_1^{20}log(1-cos(x_i-\theta))
$$





```{r}
d.2<-c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96, 2.53, 
       3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52, 2.50)

# Defind l as the log likelihood function
l<-function(t){sum(log(1-cos(d.2-t)))-20*log(2*pi)}

t.p2<-seq(-pi,pi,length.out = 2000)
y=c()

for(i in 1:2000){y[i]<-l(t.p2[i])}


plot(t.p2,y, type="l", main = "log likelihood function", ylab = "log likelihood")

```


###(b) MOM Estimator

The first moment is 
$$
E(X)=\int_0^{2\pi}(1-cos(x-\theta))/2{\pi} dx=1
$$


###(c) Newton–Raphson method

The Newton-Raphson Method is updating $\theta$ via:
$$
\theta^{(n+1)}=\theta^{(n)}-\frac{l'(\theta^{(n)})}{l''(\theta^{(n)})}
$$



```{r}
grad.l<-function(x,t){
  sum(sin(t-x)/(1-cos(t-x)))
}

hessian.2<-function(x,t){
  sum(cos(t-x)/(1-cos(t-x))-sin(t-x)^2/(1-cos(t-x))^2)
}



```




















