---
title: "Zhao Ping HW2"
author: "Ping Zhao"
date: "2/10/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

###(a) Gradient Function
```{r}
#x=(x1,x2,x3,x4), and t os the parameter (theta)
grad <- function (x, t) {
  dl = x[1]/(2+t)-(x[2]+x[3])/(1-t)+x[4]/t
}


```




###(b)Stop your algorithm when either of the following criteria is satisfied:

```{r}
secant <- function(maxit, x, t0, t1, tolerr, tolgrad){
  tstar = -1657/7680+sqrt(3728689)/7680
  digits = -log10(abs(t1 - tstar))
  it=0
  stop=0
  
  theta=c()
  converage.rate=c()
  digit=c()
  iteration=c()
  modified.relerr=c()
  Gradients=c()
  while (it < maxit & stop==0){
      it = it+1
      dl.1 = grad(x,t1)
      dl.0 = grad(x,t0)
      
      tnew = t1-dl.1*(t1-t0)/(dl.1-dl.0)
  
      
      rate = abs(tnew - tstar)/abs(t1 - tstar) #check convergent rate
      digits = -log10(abs(tnew - tstar)/abs(tstar))
      mod.relerr = abs(t1 - tnew)/max(1,abs(tnew))
      dl.new <-grad(x,tnew)
      
      if (mod.relerr < tolerr & abs(dl.new) < tolgrad) stop=1 # Stop iteration condition
      #print(c(it, t, rate,digits))
      #print(sprintf('it = %2.0f   teta = %12.12f    Rate = %4.2f     digits = %2.1f',it,tnew,rate,digits))
      #print(sprintf('          relerr = %4.1e,        grad = %4.1e',relerr,dl))
      t0 = t1 # Update t0
      t1 = tnew # Update and return
      
      theta[it]<- tnew
      converage.rate[it]<- rate
      digit[it]<-digits
      iteration[it]<- it
      modified.relerr[it] <-mod.relerr
      Gradients[it]<-dl.new
    }
  conver.info<-as.data.frame(cbind(iteration, theta, converage.rate, digit, modified.relerr, Gradients))
  return(conver.info)
}



```




### (c) &(d)

$\theta^{(0)}=.02$ and $\theta^{(1)}=.01$, and use tolerr = 1e-6 and tolgrad=1e-9

The iteration information is in matrix m. A larger matrix m.index contains information of starting points.
```{r}

d <- c(1997,907,904,32)

#secant(20,x=d,0.02,0.01,1e-6,1e-9 )


m<-secant(20,x=d,0.02,0.01,1e-6,1e-9)

# Now, I will modify the out put digit format
m.dig1<-format(m[,3:6],digit=1)
m.t<-format(m$theta, digits = 12)
m.digit<-cbind.data.frame(m$iteration,m.t,m.dig1)

names(m.digit)[1:2] <- c("iteration","theta")

m.digit
```

The matrix above gives iteration information about the secant method.  



However, I need to add information of the starting point $\theta_0$ and $\theta_1$ to the convergence matrix. As we are applying secant method, the $\theta_2$ is from the first iteration. (I don't think this index system is good.) I added an column of $\theta$ index to the matrix.

```{r}


Index.Theta<-seq(0,nrow(m)+1)
tstar = -1657/7680+sqrt(3728689)/7680

#mod.relerr.0 = abs(0.02 - tnew)/max(1,abs(tnew))
mod.relerr.1 = abs(0.01 - 0.02)/max(1,abs(0.01))

rate.t1 = abs(0.01 - tstar)/abs(0.02 - tstar)

digits.1 = -log10(abs(0.01 - tstar)/abs(tstar))
digits.0 = -log10(abs(0.02 - tstar)/abs(tstar))

t0.info<-c(NA,0.02,NA,digits.0, NA, grad(d,0.02))
t1.info<-c(NA,0.01,rate.t1,digits.1,mod.relerr.1,grad(d,0.01))

m.starting<-rbind(t0.info,t1.info,m)
m.index<-cbind(Index.Theta,m.starting)

#Now, I will modify the output digits
m.in.dig<-format(m.index[,4:7],digit=1)
m.index.digit<-cbind(m.index[,1:3],m.in.dig)


m.index.digit

```



###(e)
```{r}

rate.sq<-m$converage.rate^2

plot(m$converage.rate, ylab = "Rate", main = "Converagence Rate", pch=15)

abline(h=0, col=2)


```

As we can see from the rate plot above, the rate is less than 1 and going to 0. Thus, it is supper-linear converagence.



###(f)

$\theta^{(0)}=0.5$ and $\theta^{(1)}=0.01$

```{r}
m.f<-secant(1000,x=d,0.5,0.01,1e-6,1e-6)
plot(m.f$theta)
```

From the result above, we can see that the secant method doesn't converage in this case.

## Problem 2
(a) log likelihood function

$$
l=-20*log2\pi+\Sigma_1^{20}log(1-cos(x_i-\theta))
$$





```{r}
d.2<-c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96, 2.53, 
       3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52, 2.50)

# Defind l as the log likelihood function
l<-function(t){sum(log(1-cos(d.2-t)))-20*log(2*pi)}

t.p2<-seq(-pi,pi,length.out = 2000)
y=c()

for(i in 1:2000){y[i]<-l(t.p2[i])}


plot(t.p2,y, type="l", main = "log likelihood function", ylab = "log likelihood")

```


###(b) MOM Estimator

The first moment is 
$$
E(X)=\int_0^{2\pi}x(1-cos(x-\theta))/2{\pi} dx=sin\theta+\pi=\bar{x}
$$
Solve for MOM estimatior

```{r}
mom.est<-asin(mean(d.2)-pi)
```

###(c) Newton–Raphson method

The Newton-Raphson Method is updating $\theta$ via:
$$
\theta^{(n+1)}=\theta^{(n)}-\frac{l'(\theta^{(n)})}{l''(\theta^{(n)})}
$$



```{r}
grad.l<-function(x,t){
  sum(sin(t-x)/(1-cos(t-x)))
}

hess.2<-function(x,t){
  sum(cos(t-x)/(1-cos(t-x))-sin(t-x)^2/(1-cos(t-x))^2)
}


Newton <- function (maxit, x, t, tolerr, tolgrad) {
  stop=0
  it=0
  
  iter=c()
  Theta=c()
  Gradi=c()
  Error=c()
  while (stop==0 & it < maxit) {
    it <- it +1
    
    
    dl = grad.l(x,t)
    ddl = hess.2(x,t)
    tnew = t - dl/ddl
    dl.new<-grad.l(x,tnew)
    iter[it]<-it
    Theta[it]<-tnew
    mod.err = abs(tnew - t)/max(1,abs(tnew))
    
    
    
    iter[it]<-it
    Theta[it]<-tnew
    Gradi[it]<- dl.new
    Error[it]<-mod.err
    
    #rate = abs(tnew - tstar)/abs(t - tstar) # Not knowing tstar here

    t = tnew
    
    
    
    if(mod.err<tolerr & abs(dl.new)<tolgrad) stop=1
  } 
  
  Gradien<-format(Gradi,digits = 2)
  Errors<-format(Error,digits = 2)
  theta.12<-format(Theta, digits = 12)
  it.info<-cbind.data.frame(iter,Theta,theta.12, Gradien, Errors)
} 


```




```{r}
start.mle<-Newton(20, d.2, mom.est, 1e-6, 1e-9)
start.left<-Newton(20, d.2, -2.7, 1e-6, 1e-9)
start.right<-Newton(20, d.2, 2.7, 1e-6, 1e-9)

 

start.mle # starting point is the MOM estimator
start.left # starting point is -2.7
start.right # starting point is 2.7
```

From the outcome above, we can see that with different starting points, the $\hat\theta$ converage to different value(local optimization).

(d)
```{r}
start.sq<-seq(-pi,pi,length.out = 200)

t=c()
Gr<-c()
for(i in 1:200){
  
  t[i]<-tail(Newton(20, d.2, start.sq[i], 1e-6, 1e-9)$Theta, n=1)
  Gr[i]<-tail(Newton(20, d.2, start.sq[i], 1e-6, 1e-9)$Gradi, n=1)
}

plot.data<-cbind.data.frame(start.sq,t,Gr)



require(ggplot2)


p <- ggplot(data = plot.data, mapping = aes(x = start.sq, y = t))

p+geom_point(shape=19, alpha=0.5, color="pink")

```

From the plot above, we can see that with different starting points, the approximations given by Newton Method converge to local optimization. Also, we can see that there is a 45 degree line trend between starting points and approximation. Therefore, starting points have impact on the final convergence for Newton-Method. 

```{r}

# The following is to check different level with significant digits at 4

g <- factor(round(t, digits = 4))

xg <- split(t, g)

print(xg)

```  
  
  
From the list output, we can see that there are 19 different approximation values according to different starting points (round to 4 digit)
  
  

(e)
There are several pair of close starting points that go to different converagence. I take the 10th and 11th groups as example.

```{r}
# Check the approximation of the 10th and 11th group

xg[[10]]
xg[[11]]

```

Above are the 10th (value=-0.9533363, length of 19) the 11th group (value=-0.011972, length of 42) group for level of approximation.


```{r}
# Find the index of the flipping starting point from the 10th group to the 11th group



id<-sum(plot.data$t<= -0.9533363)
id

start.sq[id] 
t[id]
start.sq[id+1]
t[id+1]
```
  
At the 74th, the starting point is -0.8367056, with the Newton approximation -0.9533363. While for the next starting point, the 75th, the starting point value is -0.8051318, but the Newton approximation coverages to -0.011972.






